---
title: "The Modified Newton-Raphson Method"
author: "Ahmed M. Tawfik"
---

The Modified Newton-Raphson Method 
 
First, we define the function we want to use.
```{r}
fn <- function(x){x^4-8.6*x^3-35.51*x^2+464.4*x-998.46} # other function: exp(-x)-x
fn(4)
```

Second, get the derivative.
```{r}
fne<- expression(x^4-8.6*x^3-35.51*x^2+464.4*x-998.46)
fn1<- D(fne, 'x')
# this will print the derivative 
fn1 #3 * x^2 - 2 * (2 * x) + 1
```
#Third, define new function of the first derivative
```{r}
fn_1<- function(x){4 * x^3 - 8.6 * (3 * x^2) - 35.51 * (2 * x) + 464.4}
fn_1(4)

```
And define new function of the second derivative
```{r}

fne2<- expression(4 * x^3 - 8.6 * (3 * x^2) - 35.51 * (2 * x) + 464.4)
fn2<- D(fne2, 'x')
fn2

fn_2<- function(x){4 * (3 * x^2) - 8.6 * (3 * (2 * x)) - 35.51 * 2}
fn_2(4)
```
_____________________

To graph the function  we have to see the initial point
```{r}

curve(fn, from=2, to=10, xlab="x", ylab="y")
abline(h=0,col="red",lwd=1.5)
```

NOTE:  you can change the range of x (from, to) 
_____________________


Eventually we define the function of Modified Newton-Raphson 

```{r}
 
library(rlist)
ModNewRaph<- function(xi,es=0.05,j=10){ 
  i=0 #intial value for i
  ea<-1.1*es
  x0=xi
  newlist <- list()
  
  while (ea>es & i<j) { 
    x1<- x0-(fn_1(x0)*fn(x0)) / (fn_1(x0)^2-fn(x0)*fn_2(x0))
    
    ea<-abs((x1-x0)) #/x1 -> divide by x1 if you want relative error
      looplist <- list(i,x0,fn(x0),fn_1(x0),fn_2(x0),x1,ea)
      newlist <- list.append(newlist, looplist)
    
    x0=x1
    i=i+1
  }
  df <-  as.data.frame(do.call(rbind, newlist))
  list_name <- list("iteration",
                    "x(i)",
                    "f(xi)",
                    "f'(xi)",
                    "f''(xi)",
                    "x(i+1)",
                    "Error")
  names(df) <- c(list_name)
#  View(df)
  print(df)
    cat("x=",x0 ,"\n") #
}
```

```{r}
ModNewRaph(4,0.005,5)
```
```{r, echo=FALSE}
  cat("\n the following is to calculat *m*: \n \n")
```

```{r}
  
  fne <-
    expression(x ^ 4 - 8.6 * x ^ 3 - 35.51 * x ^ 2 + 464.4 * x - 998.46)
get_m <-function(x){
  fn1 <- fne
  a <- 0
  m<-0
  while (a == 0) {
    fn1 <- D(fn1, 'x')
    print(fn1)
    a <- eval(fn1)
    cat("the derivative= ",a,"\n")
    m<- m+1
    print(m)
  }
  
  #cat("x=",x0 ,"\n") #
  cat("m=",m ,"\n") #
}
get_m(4.3)
```
let's graph the function again

```{r}
 x_1 = 4.3
 curve(
   fn,
   from = 3,
   to = 8,
   xlab = "x",
   ylab = "y"
 )
 abline(h = 0, col = "red", lwd = 1.5)
 abline(v = x_1, col = "red" , lwd = 1.5)
 text(x = 5, y = 15, labels = "(4.3,0)")
```

rmarkdown::render("Modified Newten.R")

